---
title: "SC_QC_Ambient_Doublet_Removal"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
---

**R v4.4.1 & Seurat v5.1**
**Open the project "cm_wd".** 

Code is based on Jonathan's "IRI12h1_1_Ambient RNA and Doublet Removal_template.Rmd". Run 6 samples using this format. Each run result is saved as sample_name.html, and each final (cleaned) seurat object is saved in results directory.

<br>

### Where I am

```{r}
#Sanity check where here is located
here::here()
```

### Set working environment

```{r}
# load libraries from 00_project_setup 
source(
  fs::path(here::here(), 
           "00_project_setup", 
           "!libraries.r")
  )

# load directory path from 00_project_setup 
source(
  fs::path(here::here(), 
           "00_project_setup", 
           "!directories.r")
  )
```


### Data directory

```{r}
dir_cellranger_out
```

### Results directory (to save seurat objects)

```{r}
dir_01_seurat_objects
```

### Sample info

```{r results=FALSE}

library(readr)
sample_info <- read_csv(fs::path(dir_data, 
                                "cellranger_output",
                                "sample_info.csv"))

sample_info
```

<br><br><br>

# Sample & Condition

```{r}
sample.num <- 6   # change

Sample <- paste(sample_info[sample.num,1])
Sample
Condition <- paste(sample_info[sample.num,2])
Condition
```

<br>

# SoupX for Ambient RNA Removal

[(Young and Behjati, 2020)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7763177/)

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
# Load data and estimate soup profile
tod = Read10X_h5(fs::path(dir_cellranger_out, Sample, "/raw_feature_bc_matrix.h5")) #Change
toc = Read10X_h5(fs::path(dir_cellranger_out, Sample, "/filtered_feature_bc_matrix.h5")) #Change
sc = SoupChannel(tod,toc)

#Make the Seurat object from the filtered control data
SO <- Read10X_h5(fs::path(dir_cellranger_out, Sample, "/filtered_feature_bc_matrix.h5"))#Change 
SO <- CreateSeuratObject(counts = SO, project = "PFAS_kidney_organoid")  #Change
 
#Cluster the cells with Seurat
SO <- SCTransform(SO, verbose = F)
SO <- RunPCA(SO, verbose = F)
SO <- RunUMAP(SO, dims = 1:30, verbose = F)
SO <- FindNeighbors(SO, dims = 1:30, verbose = F)
SO <- FindClusters(SO, verbose = T)
 
meta <- SO@meta.data
umap <- SO@reductions$umap@cell.embeddings
clusters <- setNames(meta$seurat_clusters, rownames(meta))

#Sanity Check
length(clusters) #should be equal to nrow(sc$metaData)
nrow(sc$metaData)

sc <- setClusters(sc, clusters)
sc <- setDR(sc, umap)

#Estimate rho
sc = autoEstCont(sc)

#Clean the data
SO_out = adjustCounts(sc)
 
#Create a new Seurat Object out of the cleaned data
seurat.obj <- CreateSeuratObject(SO_out)
```

<br>

# Pre-process for Doublet Finder

[(McGinnis et al., 2019)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6853612/)

```{r echo=TRUE, error=FALSE, fig.align= 'center', message=FALSE, warning=FALSE}

seurat.obj

seurat.obj[["percent.mt"]] <- PercentageFeatureSet(seurat.obj, pattern = "^MT-")

VlnPlot(seurat.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r echo=TRUE, error=FALSE, fig.align= 'center', message=FALSE, warning=FALSE}
#Set lower filter to get rid of low count nFeature & MT %
seurat.obj.f <- subset(seurat.obj, subset = nFeature_RNA > 300)
seurat.obj.f <- subset(seurat.obj.f, subset = percent.mt < 20)


VlnPlot(seurat.obj.f, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

seurat.obj.f
```

```{r echo=TRUE, error=FALSE, fig.align= 'center', message=FALSE, warning=FALSE}
# Pre-process standard workflow
seurat.obj.f <- NormalizeData(object = seurat.obj.f)
seurat.obj.f <- FindVariableFeatures(object = seurat.obj.f)
seurat.obj.f <- ScaleData(object = seurat.obj.f)
seurat.obj.f <- RunPCA(object = seurat.obj.f)
ElbowPlot(seurat.obj.f, ndims = 40)
```

```{r echo=TRUE, error=FALSE, fig.align= 'center', message=FALSE, warning=FALSE}
# PCs between 15-20
seurat.obj.f <- FindNeighbors(object = seurat.obj.f, dims = 1:30)

seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.03, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.05, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.1, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.2, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.3, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.4, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.5, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.6, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.7, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.8, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.9, verbose = FALSE)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 1, verbose = FALSE)


seurat.obj.f <- RunUMAP(object = seurat.obj.f, dims = 1:30)
```

```{r}
# check resolution
DimPlot(seurat.obj.f, reduction = "umap", group.by = "RNA_snn_res.0.03")
DimPlot(seurat.obj.f, reduction = "umap", group.by = "RNA_snn_res.0.05")
DimPlot(seurat.obj.f, reduction = "umap", group.by = "RNA_snn_res.0.1")
DimPlot(seurat.obj.f, reduction = "umap", group.by = "RNA_snn_res.0.2")
DimPlot(seurat.obj.f, reduction = "umap", group.by = "RNA_snn_res.0.3")
```


```{r echo=TRUE, error=FALSE, fig.align= 'center', message=FALSE, warning=FALSE, fig.asp=0.4}

# tentative setting
seurat.obj.f$seurat_clusters <- seurat.obj.f$RNA_snn_res.0.2  # change

p1 <- DimPlot(seurat.obj.f, reduction = "umap", group.by = "seurat_clusters", label = TRUE) + NoLegend()
p2 <- VlnPlot(seurat.obj.f, features = c("nFeature_RNA"), group.by = "seurat_clusters")

p1+p2
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', results=F}

# Calculate each combination of pN and pK
sweep.res.list_seurat.obj.f <- paramSweep(seurat.obj.f, PCs = 1:20, sct = FALSE) 
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

#Summarize each combination of pN and pK
sweep.stats_seurat.obj.f <- summarizeSweep(sweep.res.list_seurat.obj.f, GT = FALSE) 

#Select the pK that corresponds to max bcmvn to optimize doublet detection
bcmvn_seurat.obj.f <- find.pK(sweep.stats_seurat.obj.f)
pK <- bcmvn_seurat.obj.f %>% 
  filter(BCmetric == max(BCmetric)) %>%
  dplyr::select(pK) 

#See pK in the Values Environment
pK <- as.numeric(as.character(pK[[1]]))

```

```{r homotypic_doublets, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
 
# Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
annotations <- seurat.obj.f@meta.data$seurat_clusters  
 
homotypic.prop <- modelHomotypic(annotations)           
homotypic.prop
 
# 10X Multiplet Rate Table (the doublet ratio is # of cells recovered divided by 125000) https://kb.10xgenomics.com/hc/en-us/articles/360001378811-What-is-the-maximum-number-of-cells-that-can-be-profiled-
 
nrow(seurat.obj.f@meta.data)

nExp_poi <- round(nrow(seurat.obj.f@meta.data) # To calculate cell number
                  /125000              # To calculate the doublet ratio
                  *nrow(seurat.obj.f@meta.data))
nExp_poi

#nExp_poi <- 1000

nExp_poi_adj <- round(nExp_poi*(1-homotypic.prop))

nExp_poi_adj 
```

<br>

# Doublet Finder

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

seurat.obj.f_doublets <- doubletFinder(seurat.obj.f,
                        PCs = 1:20,
                        pN = 0.25,
                        pK = pK,
                        nExp = nExp_poi_adj,
                        reuse.pANN = FALSE, sct = FALSE)
```

```{r}
colnames(seurat.obj.f_doublets@meta.data)[19] <- "pANN"  # change
colnames(seurat.obj.f_doublets@meta.data)[20] <- "DF.class"  # change
head(seurat.obj.f_doublets@meta.data)
table(seurat.obj.f_doublets@meta.data$DF.class)

DimPlot(seurat.obj.f_doublets, group.by = "DF.class")

```

<br>

# Subset singlets

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

seurat.obj.f_singlets <- subset(seurat.obj.f_doublets, DF.class == "Singlet")
seurat.obj.f_singlets
DimPlot(seurat.obj.f_singlets, reduction = "umap", group.by = "seurat_clusters")

```

<br>

# Add Meta data

```{r echo=TRUE, error=FALSE, fig.align= 'center', message=FALSE, warning=FALSE}
# Create Seurat object and add metadata

seurat.obj.f_singlets <- AddMetaData(object = seurat.obj.f_singlets, metadata = Sample, col.name = "Sample") #Change
seurat.obj.f_singlets <- AddMetaData(object = seurat.obj.f_singlets, metadata = Condition, col.name = "Condition") #Change


head(seurat.obj.f_singlets@meta.data)
```

<br>

# Set Idents

```{r}
Idents(seurat.obj.f_singlets) <- "seurat_clusters"
```

<br>

# Initial Cluster Identification with Heatmap and Vln Plot

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

seurat.obj.f_singlets.markers <- FindAllMarkers(seurat.obj.f_singlets, 
                                                only.pos = TRUE, 
                                                min.pct = 0.25, 
                                                logfc.threshold = 0.25,
                                                verbose = FALSE)
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width = 7, fig.height= 8}
# heatmap with top5
seurat.obj.f_singlets.markers %>%
    group_by(cluster) %>%
    top_n(n = 5, wt = avg_log2FC) -> top5

DoHeatmap(seurat.obj.f_singlets, features = top5$gene) + NoLegend()

# violin plot with top2
seurat.obj.f_singlets.markers %>%
    group_by(cluster) %>%
    top_n(n = 2, wt = avg_log2FC) -> top2

VlnPlot(seurat.obj.f_singlets, 
        features = unique(top2$gene),
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0)+
        NoLegend()

# list of top10
t1 <- seurat.obj.f_singlets.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10)
   
t1
```

<br>

# Initial Cluster Identification with DotPlot

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DotPlot(seurat.obj.f_singlets, features = unique(top2$gene), dot.scale = 8, cols = c("dodgerblue2", "coral2")) + RotatedAxis()
```

<br>

# Code to Add to Pre-processing

```{r fig.asp=0.5}
# nFeature check

p1 <- DimPlot(seurat.obj.f_singlets, reduction = "umap", label = TRUE) + NoLegend()
p2 <- FeaturePlot(seurat.obj.f_singlets, features = "nFeature_RNA")
p3 <- VlnPlot(seurat.obj.f_singlets, features = "nFeature_RNA")

p1+p2
p3
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=TRUE, fig.align = 'center', fig.asp=0.5}
# Jonathan's cell type marker set
cluster <- c("COL1A1",    # for Fibroblasts
             "VCAM1",     # for Damaged Epithelial
             "EPCAM",     # for Epithelial
             "GATA3",     # for Mesenchymal Unknown
             "TOP2A",     # for Cycling
             "PIEZO2")    # for Mesangial Cells

for (i in cluster) {
  p1 <- FeaturePlot(seurat.obj.f_singlets, i, order = T)
  p2 <- VlnPlot(seurat.obj.f_singlets, i) + NoLegend()
  #print(p1)
  #print(p2)
  print(p1|p2)
}
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=TRUE, fig.align = 'center'}

DotPlot(seurat.obj.f_singlets,
features = cluster,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5) +
  coord_flip() +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        text = element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab(NULL) +
  ylab(NULL)
```

<br>

# Renaming Clusters and Re-Run Markers

```{r echo=TRUE, warning=FALSE, error=FALSE, message=TRUE, fig.align = 'center', fig.width=8}

seurat.obj.f_singlets@meta.data <- seurat.obj.f_singlets@meta.data %>% 
  mutate(class.jwn = dplyr::case_when(
    seurat_clusters == 0  ~ "Damaged Epithelial",
    seurat_clusters == 1  ~ "Fibroblasts",
    seurat_clusters == 2  ~ "Fibroblasts",
    seurat_clusters == 3  ~ "Fibroblasts",
    seurat_clusters == 4  ~ "Epithelial",
    seurat_clusters == 5  ~ "Unknown Cells",
    seurat_clusters == 6  ~ "Mesenchymal Unknown",
    seurat_clusters == 7  ~ "Cycling",
    seurat_clusters == 8  ~ "Mesangial Cells"

 ))

# set an order 
seurat.obj.f_singlets$class.jwn <- factor(seurat.obj.f_singlets$class.jwn, 
                                          levels = c("Fibroblasts", 
                                                     "Damaged Epithelial", 
                                                     "Epithelial",
                                                     "Mesenchymal Unknown",
                                                     "Cycling",
                                                     "Mesangial Cells",
                                                     "Unknown Cells"))

table(seurat.obj.f_singlets$class.jwn)
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=TRUE, fig.align = 'center', fig.width=8}
Idents(seurat.obj.f_singlets) <- seurat.obj.f_singlets@meta.data$class.jwn

DimPlot(seurat.obj.f_singlets, group.by = "class.jwn") + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=15))

DimPlot(seurat.obj.f_singlets, group.by = "class.jwn", label = T) + ggtitle(NULL) + xlab("UMAP 1") + ylab("UMAP 2") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        text = element_text(size=15))

```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=TRUE, fig.align = 'center', fig.asp=0.6}

for (i in cluster) {
  p1 <- FeaturePlot(seurat.obj.f_singlets, i, order = T)
  p2 <- VlnPlot(seurat.obj.f_singlets, i) + NoLegend()
  #print(p1)
  #print(p2)
  print(p1|p2)
}
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=TRUE, fig.align = 'center', fig.height=6}
DotPlot(seurat.obj.f_singlets,
features = cluster,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5) +
  coord_flip() +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        text = element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab(NULL) +
  ylab(NULL)
```

<br>

# Output Files

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

saveRDS(seurat.obj.f_singlets, paste0(dir_01_seurat_objects, "/", Sample, ".rds")) #Change

```

<br>

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()
```
