---
title: "SC_PFNA_DEG"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
---

**R v4.4.1 & Seurat v5.1**
**Open the project "cm_wd".** 

<br>

### Where I am

```{r}
#Sanity check where here is located
here::here()
```

### Set working environment

```{r results=FALSE}
# load libraries from 00_project_setup 
source(
  fs::path(here::here(), 
           "00_project_setup", 
           "!libraries.r")
  )

# load directory path from 00_project_setup 
source(
  fs::path(here::here(), 
           "00_project_setup", 
           "!directories.r")
  )
```


### Data directory

```{r}
dir_06_seurat_objects
```

### Results directory

```{r}
# for seurat object
dir_07_seurat_objects

# for results
dir_07_results
```

### Sample info

```{r results=FALSE}

library(readr)
sample_info <- read_csv(fs::path(dir_data, 
                                "cellranger_output",
                                "sample_info.csv"))

sample_info
```

<br><br><br>

## Load merged.filtered data
```{r}
ks.merged.filtered <- readRDS(paste(fs::path(dir_06_seurat_objects), "/ks.PFNA.merged.filtered.rds", sep = ""))
```

<br>

## Join data
```{r}
SO <- JoinLayers(ks.merged.filtered)
```

```{r}
ks.merged.filtered

SO
```

```{r}
head(SO@meta.data)
tail(SO@meta.data)
```

<br> 

## Add Metadata
```{r}
SO@meta.data <- SO@meta.data %>%
  mutate(class.PFAS = dplyr::case_when(
    Sample == "KS-01"  ~ "Veh",
    Sample == "KS-02"  ~ "PFNA",
    Sample == "KS-04"  ~ "Veh",
    Sample == "KS-05"  ~ "PFNA",
  ))
 
SO@meta.data <- SO@meta.data %>%
  mutate(class.time = dplyr::case_when(
    Sample == "KS-01"  ~ "1_Week",
    Sample == "KS-02"  ~ "1_Week",
    Sample == "KS-04"  ~ "2_Weeks",
    Sample == "KS-05"  ~ "2_Weeks",
  ))

# set an order in Condition
SO$Condition <- factor(SO$Condition, 
                       levels = c("CTR_wk1","PFNA_wk1","CTR_wk2","PFNA_wk2"))
```

<br>

```{r}
table(SO$Condition)
```

<br>

```{r fig.asp=0.6}
p1 <- DimPlot(SO, group.by = "cell_type", label = TRUE) + NoLegend()
p2 <- DimPlot(SO, group.by = "Condition")

p1+p2
```

<br>

## Save seurat object
```{r}
# save object
saveRDS(SO, paste0(dir_07_seurat_objects, "/ks.PFNA.merged.filtered.joined.rds"))
```


<br><br>

## Proportion plots {.tabset}

### By Sample
```{r}
# set cell type order
cluster_order <- match(levels(SO@meta.data[["cell_type"]]), metaLevels("cell_type", SO))

dittoSeq::dittoBarPlot(SO, var = "cell_type", group.by = "Sample", var.labels.reorder = cluster_order)
```

### By Duration
```{r}
dittoSeq::dittoBarPlot(SO, var = "cell_type", group.by = "class.time", var.labels.reorder = cluster_order)
```

### By Treatment
```{r}
dittoSeq::dittoBarPlot(SO, var = "cell_type", group.by = "class.PFAS", var.labels.reorder = cluster_order)
```

##

<br>

## Cluster markers
```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
Idents(SO) <- "cell_type"
KS.cluster.markers <- FindAllMarkers(SO, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, verbose = FALSE)
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width = 7, fig.height= 8}
# heatmap with top5
KS.cluster.markers %>%
    group_by(cluster) %>%
    top_n(n = 5, wt = avg_log2FC) -> top5


DoHeatmap(SO, features = top5$gene) + NoLegend()

# violin plot with top2
KS.cluster.markers %>%
    group_by(cluster) %>%
    top_n(n = 2, wt = avg_log2FC) -> top2

VlnPlot(SO, 
        features = unique(top2$gene),
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0)+
        NoLegend()

# list of top10
t1 <- KS.cluster.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10)
   
t1

DoHeatmap(SO, features = t1$gene) + NoLegend()

```




<br><br><br>

## Comparison
```
Week1_Veh vs. Week1_PFNA for each cluster
Week2_Veh vs. Week2_PFNA for each cluster
Week1_Veh vs. Week2_Veh for each cluster
``` 


## 1. Week1_PFNA vs. Week1_Veh 

```{r}
cluster.final <- unique(SO$cell_type)

for (i in cluster.final) {

## DE markers per class.jwn.2 (between samples) ##   
  # subset by final cluster annotation "i" cells from data.integrated
  Idents(SO) <- "cell_type"
  cluster.subset <- subset(SO, idents = i)
  
  # find marker between samples
  Idents(cluster.subset) <- "Condition"
  de.markers <- FindMarkers(cluster.subset, 
                            ident.1 = "PFNA_wk1",       # Case
                            ident.2 = "CTR_wk1",        # Control  
                            logfc.threshold = .25, min.pct = .1)   # Jonathans setup  
  
  de.markers <- de.markers[order(de.markers[,"avg_log2FC"], decreasing = TRUE),]  # Jonathans setup  

  assign(paste0("PFNAvsCTR_week1_",i,".de.markers"), de.markers)
  
  write.csv(de.markers, file = paste0(dir_07_results, "/PFNAvsCTR_week1_", i, ".de.markers.csv"))
  
}
```

### DEG list {.tabset}
```
Positive logFC: High expression in case 
Negative logFC: High expression in control
```

#### Epithelial_1
```{r}
datatable(`PFNAvsCTR_week1_Epithelial_1.de.markers`)
```

#### Epithelial_2
```{r}
datatable(PFNAvsCTR_week1_Epithelial_2.de.markers)
```

#### Fibroblasts
```{r}
datatable(`PFNAvsCTR_week1_Fibroblasts.de.markers`)
```

#### Mesangial_Cells_1
```{r}
datatable(`PFNAvsCTR_week1_Mesangial_Cells_1.de.markers`)
```

#### Mesangial_Cells_2
```{r}
datatable(`PFNAvsCTR_week1_Mesangial_Cells_2.de.markers`)
```

#### Cycling
```{r}
datatable(PFNAvsCTR_week1_Cycling.de.markers)
```

##

<br>

## 2. Week2_PFNA vs. Week2_Veh 

```{r}
cluster.final <- unique(SO$cell_type)

for (i in cluster.final) {

## DE markers per class.jwn.2 (between samples) ##   
  # subset by final cluster annotation "i" cells from data.integrated
  Idents(SO) <- "cell_type"
  cluster.subset <- subset(SO, idents = i)
  
  # find marker between samples
  Idents(cluster.subset) <- "Condition"
  de.markers <- FindMarkers(cluster.subset, 
                            ident.1 = "PFNA_wk2",       # Case
                            ident.2 = "CTR_wk2",        # Control  
                            logfc.threshold = .25, min.pct = .1)   # Jonathans setup
  
  de.markers <- de.markers[order(de.markers[,"avg_log2FC"], decreasing = TRUE),]  # Jonathans setup  

  assign(paste0("PFNAvsCTR_week2_",i,".de.markers"), de.markers)
  
  write.csv(de.markers, file = paste0(dir_07_results, "/PFNAvsCTR_week2_", i, ".de.markers.csv"))
  
}
```

### DEG list {.tabset}
```
Positive logFC: High expression in case 
Negative logFC: High expression in control
```

#### Epithelial_1
```{r}
datatable(`PFNAvsCTR_week2_Epithelial_1.de.markers`)
```

#### Epithelial_2
```{r}
datatable(PFNAvsCTR_week2_Epithelial_2.de.markers)
```

#### Fibroblasts
```{r}
datatable(PFNAvsCTR_week2_Fibroblasts.de.markers)
```

#### Mesangial_Cells_1
```{r}
datatable(`PFNAvsCTR_week2_Mesangial_Cells_1.de.markers`)
```

#### Mesangial_Cells_2
```{r}
datatable(`PFNAvsCTR_week2_Mesangial_Cells_2.de.markers`)
```

#### Cycling
```{r}
datatable(PFNAvsCTR_week2_Cycling.de.markers)
```

##

<br>

## 3. Week2_Veh  vs. Week1_Veh 

```{r}
cluster.final <- unique(SO$cell_type)

for (i in cluster.final) {

## DE markers per class.jwn.2 (between samples) ##   
  # subset by final cluster annotation "i" cells from data.integrated
  Idents(SO) <- "cell_type"
  cluster.subset <- subset(SO, idents = i)
  
  # find marker between samples
  Idents(cluster.subset) <- "Condition"
  de.markers <- FindMarkers(cluster.subset, 
                            ident.1 = "CTR_wk2",       # Case
                            ident.2 = "CTR_wk1",        # Control  
                            logfc.threshold = .25, min.pct = .1)   # Jonathans setup  
  
  de.markers <- de.markers[order(de.markers[,"avg_log2FC"], decreasing = TRUE),]  # Jonathans setup  
  
  assign(paste0("CTR_week2vsCTR_week1_",i,".de.markers"), de.markers)
  
  write.csv(de.markers, file = paste0(dir_07_results, "/CTR_week2vsCTR_week1_", i, ".de.markers.csv"))
  
}
```

### DEG list {.tabset}
```
Positive logFC: High expression in case 
Negative logFC: High expression in control
```

#### Epithelial_1
```{r}
datatable(`CTR_week2vsCTR_week1_Epithelial_1.de.markers`)
```

#### Epithelial_2
```{r}
datatable(CTR_week2vsCTR_week1_Epithelial_2.de.markers)
```

#### Fibroblasts
```{r}
datatable(CTR_week2vsCTR_week1_Fibroblasts.de.markers)
```

#### Mesangial_Cells_1
```{r}
datatable(`CTR_week2vsCTR_week1_Mesangial_Cells_1.de.markers`)
```

#### Mesangial_Cells_2
```{r}
datatable(`CTR_week2vsCTR_week1_Mesangial_Cells_2.de.markers`)
```

#### Cycling
```{r}
datatable(CTR_week2vsCTR_week1_Cycling.de.markers)
```

##

<br>

## Session info
```{r}
sessionInfo()
```



